### 2015-05-10 thesis/thesis/chap5/chap5.tex
added coop

\chapter{LEarning Chapter}\label{cha:PCTLKC+}

In this chapter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Summary}\label{sec:conclusion}

In this chapter,

>>>>

\setcounter{chapter}{4}

\chapter{Training Chapter}\label{cha:PCTLKC}

\section{Introduction}

Over the past years, online services have become an important part of many scalable business applications. The increasing reliance on online service providers has significantly influenced the way web services are engineered. Therefore, given the dynamic and unpredictable nature of the Internet, delivering high quality services is still a critical and challenging issue. One practical solution towards delivering high quality services is utilizing intelligent decision making agents. These agents aim at maximizing their gain by exploring the best ways to provide services that satisfy end users\cite{Zeng:2003:QDW:775152.775211, 10.1109/ARES.2008.7, Demirkan2013412, journals/tsc/ZhengZYB13, Josang:2007:STR:1225318.1225716}. However, agent-based web services are functionally limited in the sense that they cannot handle a large number of requests at the same time without compromising the quality of service provided. Recent developments have attempted to shift web services from passive models, consisting of individual components, to models made up of autonomous and group-based components that share common goals. In group-based models, interaction, composition, and cooperation are the key challenges \cite{ICWS2011-1, SCC2011-1, journals/mags/BaldoniBM10, journals/jcss/CasadoYT13} that directly impact the group's overall performance in achieving common goals. To that end, we see the emergence of web service \emph{communities}, which consist of services with similar functionalities but distinct nonfunctional properties \cite{Zeng:2003:QDW:775152.775211, Paik:2005:TSS:2229263.2230038, Medjahed05adynamic, 10.1109/ARES.2008.7}. A \emph{community} of web services runs continuous performance assessment functions that regulate web services' interactions and manage their composition and cooperation.

The advantages of community of web services are that the community facilitates the discovery of web services and provides better quality of service compared to passive models with individual components. Communities act as abstract web services, communicating with external entities via the same standard protocols that a normal web service employs. But, they regulate the service process via sophisticated internal communication protocols, thereby providing services based on the combined work of a number of web services. The downside to communities is the complexity of management involved in finding and inviting individual web services and managing the overall quality of the combined work of several web services.
% because although they have similar functionality, they have different attitudes.
When interacting with communities of web services, users send their requests to the coordinators of the community, which play the role of community representative or access point. The community coordinator is responsible of receiving tasks and delivering services. Moreover, as community representative, it verifies the credentials of new web services before accepting them into the community and kicks web services that could harm the value of the community.

\textbf{Challenge and Problem Statement.} In recent work, communities of web services have been proposed in order to facilitate discovery of web services, improve the Quality of Service (QoS), and help individual services find better market share and opportunities. However, there is a twofold challenge that yet to be addressed. The first part is community development, which involves finding the right web services for the community, satisfying the objectives of all the web services and community coordinators involved within the community; and the second part is from the individual web service's side because each web service aims to join the community that would guarantee the best outcome. The advocated solutions \cite{conf/webist/MaamarLBTS07, 10.1109/ARES.2008.7, journals/soca/XuYLZB11, 10.1109/TSC.2012.12, managing-hela-jalel, DBLP:conf/IEEEscc/KhosravifarABT11, DBLP:conf/IEEEscc/LimTMB12} have attempted to address these research problems. These research problems are reflected by the two challenge parts:
\begin{enumerate}
	\item The solutions consider the architecture of centralized management for communities where most of the decisions are made by the centralized coordinator. The problem is that in real world scenarios, decisions made by independent service providers are highly distributed.
	\item The solutions either propose complex algorithms \cite{DBLP:conf/IEEEscc/LimTMB12, 10.1109/TSC.2012.12, 10.1109/TSC.2014.2312940} to find the optimal strategy to follow, or oversimplify the problem by eliminating important parameters and using approximation techniques to make the algorithms tractable \cite{10.1109/TSC.2012.12}.
\end{enumerate}
These approximation methods sometimes negatively influence the outcome because simplifying the constraints may cause important aspects of the problem to be ignored.

For example, instead of calculating the gain distribution using the complex shapely-value method, \cite{10.1109/TSC.2012.12} proposes a simple egalitarian way of distributing gain, which completely ignores the gain generated from collaborative work of sub-communities. Other categories of related work exist that do not focus on all the parties involved within the community \cite{DBLP:conf/IEEEscc/KhosravifarABT11}.
%And some other work which do not focus on rationality on web services or communities involved [refrences].
In \cite{journal-community-formation} we proposed a cooperative game theory based model for aggregating web services in communities. A centralized decision maker in communities based on a complete knowledge of available web service quality metrics and performance, has been used to form optimal and stable communities that maximize individual and group income. However, centrality and complete information are strong assumptions, which are not very compatible with real business scenarios.

\textbf{Contributions.} In this paper, we introduce DDM, a Distributed Decision Making model for  community formation that regulates web service agents’ decision making process in terms of cooperating and deciding which group to join and which service to invite for joining. The decision model in extracted from a data model in the form of information obtained from a large number of web services regarding their single and cooperative utilities as well as environmental parameters such as demand, service quality, etc. The generated decision tree improves agents' understanding of the environment and how to select actions that lead towards maximizing their utilities. The advantage of this approach is that the tree is created from the past data and reflects comprehensive vision about agents' attitude of action selection. However, the tree is subject to get continuously modified to adapt with environmental changes.
%The training model deploys a logistic regression algorithm to build a hypothesis function that can thoroughly address the aforementioned research problems.
The decision model provides web services with enough information which helps those services to efficiently decide and predict the outcome of their different possible collaborations. This model works in a distributed manner in which services are self-sufficient in decision making process and do not rely on a centralized decision making process. Our findings show that communities of web services can efficiently find the appropriate web service to invite for cooperation as well as allowing a single web service to find the best communities to join. In other words, the web services agents are equipped with a recommender system to suggest beneficial action. Moreover, communities can consider the decision model and analyze the characteristics of different individual web services and make prudent decisions when inviting a web service to join or accepting a join inquiry initiated from a web service. In general, DDM equips web services with efficient methods for foreseeing how their choices will impact their short-term and long-term goals; therefore, opting for best decision available.

To effectively generate the decision model for web services, we used real data set to extract web services' individual characteristics and used them to measure outcomes when they cooperate with one another. The data set has been extracted by real-world QoS evaluation results from 142 users on 4,532 Web services on 64 different time slots. Combining these data based on each web service point of view on different time slots, we acquired 5 different unique features for 4,532 web services. By engineering and extracting these features, we gathered functional and cooperative features for both individual web services and communities in different time slots. We were able to investigate the path a web service might take to achieve the best utility out of effective interaction with others. All paths and the outcomes are labeled to be utilized in the training model. Using cross validation sets, web services are able to compute the optimal hypothesis function (using logistic regression) that can be used to predict outcomes of cooperative work with other individual web services or communities. Our findings show that web services equipped with DDM have by far better outcomes in the form of cooperative service providers than the ones that either do not cooperate or randomly find communities to join.

\section{Preliminaries and Challenging Issues}\label{s4:preliminaries}
%In this section, we first present the architecture of DDM. We explore the characteristics of intelligent service agents and the features we extract for training. To do this, we first discuss some preliminaries.
In this section, we discuss the preliminary concepts of communities of web service and we introduce the challenges of community formation.

\subsection{Web Services}\label{s:ws}

In recent years, online services have become a standard part of daily life around the globe. Many modern applications rely on web services from different providers. For instance, many mobile and tablet applications that have limited storage and processing power are merely aggregating information from different online services. Examples are vast, including weather forecasting, ticket selling, shopping apps, local maps and location services.

The World Wide Web Consortium (W3C) defines web services as ``software systems designed to support interpretable machine-to-machine interaction over a network. It has an interface described in a machine-processable format (specifically WSDL). Other systems interact with the web service in a manner prescribed by its description using SOAP messages, typically conveyed using HTTP with XML serialization in conjunction with other Web-related standards.'' When developers declare a new web service, it will be
discovered based on its description, which fully discloses its functionalities. Developers also have to declare a public interface and a readable documentation to help other developers when integrating different services \cite{w3cwsdl}. Nowadays, web API, standards that do not require XML-based web service protocols like SOAP and WSDL are also emerging. They are called REST (representational state transfer) services, which are moving towards simpler communication protocols.
%They are not restricted
%to XML formats, recently JSON, a human readable and simpler format
%is becoming popular among online service providers.

We are not going to delve into the engineering details of online web service implementation and its protocols in this paper. We are interested in web services from a business model perspective. Service providers usually charge end users for services they provide. For example, Google has listed pricing and plans for a wide range of services they provide on their web service console page\footnote{https://code.google.com/apis/console}.

In this paper, we abstract web services as rational agents\footnote{The term rational is used here in the sense that web services are utility maximizers} that provide services to end users. They aim to maximize their individual income  by receiving enough requests from end users. In order to increase their revenues, web services seek for more tasks if they have the capacity and throughput to do so. Web services can join communities to enhance efficiency by collaborating with others, to have access to broad market share, and for the opportunity to receive a bigger task pool from end users. Furthermore, the high reliance on web services has resulted in increased quality expectations from end users. Communities of web services can provide higher availability, performance, reliability, and recovery for end users.

\subsection{Web Service Communities}\label{s:wsc}

The community of web services is essentially a virtual group of web services having similar and complementary functionalities\cite{DBLP:journals/ijebr/MaamarSTBB09}. Communities aggregate web services and communicate with other entities such as UDDI registries and users, using identical protocols to those used by single web services. Web services join communities to increase utility by having a larger market share and task pool. Community coordinators are responsible for community development, managing membership requests from web services and distributing user tasks among the community members. Community coordinators try to attract quality web services to join and keep the community as stable and productive as possible to gain a better reputation and user satisfaction, which increases the community's market share. How web services reside within communities and how communities of web services are engineered is described comprehensively in \cite{DBLP:journals/ijebr/MaamarSTBB09}.

\subsection{The Join Challenge}\label{s:tjc}
Sometimes, web services can increase their overall utility by collaborating with other web services as communities. This collaboration helps them with better ways of sharing resources and having higher reputation, better market share and wider visibility. Web services and communities come with different internal and external parameters, and the long-term outcome is dependent on these parameters.

The goal of all parties involved in the community is to maximize their long-term outcome while they are operating as part of the community. Web services need to be equipped with a selection strategy to choose from all of the different possible collaboration groups they can form as well as an estimation method for evaluating the long-term gain of joining different possible communities. Web services need to experiment with different possible collaborative groups and other web services in order to estimate their gain over time. However, with a high number of possible communities, it is not possible to test collaboration with random web services. Even if a linear approximate function for estimating utility based on community web services' parameters is adopted, the exponential \footnote{Bell number: $http://en.wikipedia.org/wiki/Bell_number$} growth rate of the possible number of partitions of web services into communities would make any brute-force type algorithm for the best community selection strategy intractable and impractical in real-world application settings.

\subsection{Join Consequences}\label{s:jc}
It is worth mentioning that a \emph{join} event takes place as a result of two parties that are looking to expand their collaborations. All actions are chosen in an attempt to enhance the overall outcome. However, the selected action may result in decreasing the overall utility in the long run.
% long term.
This is the case when a single web service joins a community, but the complex process of task allocation eliminates the visibility of the web service and, thus, the web service stays idle within the community, which makes the join action a bad decision. The same event might be beneficial for the community, as it hosts a new web service that can engage in performing a new coming task. But overall, neither side benefits from collaborating with the other and the join event has negative consequences for at least one side's utility.

The more common scenario is when both parties benefit from the joining of a web service to a community. This action is rational, as both the web service and the community enhance the utility. However, the community may not be the best choice for the web service to join. In other words, the web service could have joined a better community, but, due to of lack of knowledge about the surrounding environment, joined the available community which was not the optimal choice. Since the community does enhance its utility, the web service could stay with that community, which results in a sub-optimal increase in web service's utility. In the following section the proposed model provides solutions that effectively address the aforementioned challenges.

\section{The Model Components}\label{s:themodelcomponents}

In this section, we discuss the parameters that we use in the rest of the paper. Then we present the task distribution and revenue model of our distributed web services communities.

\subsection{Internal Features}\label{s:if}

With a group of web services having identical or similar functionalities, QoS metrics provide nonfunctional characteristics for optimal candidate selection. Web service quality metrics were presented in \cite{Ardagna:2007:ASC:1263152.1263531,Menasce:2002:QIW:613357.613758,10.1109/ISSRE.2011.17}. We have adopted the most representative QoS properties of web services that highly influence the utility of web services.
%We refer to a typical web service as ws_{i}.

Let $C = \{ws_1,ws_2,..., ws_n\}$ be a community with $|n|$ web services. We define the following features for the group of web services based on their functional parameters:

\begin{itemize}

  \item \emph{Throughput $(Th_{C})$} is the rate at which a service can process requests. QoS measures can include the maximum throughput or a function that describes how throughput varies with load intensity. Throughput is a positive real number. For communities the expected throughput value of collaborating web services can be estimated as summation of throughputs of all the services inside:

	\begin{equation}
		 Th_{C} = \left\{
			\begin{array}{l l}
				Th_{w} & \quad \text{if $|C| = 1$}\\
				\sum_{w \in C}{(Th_{w})} & \quad \text{if $|C| > 1$}
			\end{array} \right.
	\end{equation}

	\item \emph{Availability $(A_{C})$} is the percentage of time that a service is operating.
	The probability that the service operation is accessible. Availability is also a real number in the range $[0, 1]$. For communities the expected availability of collaborating web services considering they follow a parallel system can be estimated as:

	\begin{equation}
		A_{C} = \left\{
			\begin{array}{l l}
				A_{w} & \quad \text{if $|C| = 1$}\\
				1-\prod_{w \in C}{(1-A_{w})} & \quad \text{if $|C| > 1$}
			\end{array} \right.
	\end{equation}

	\item \emph{Execution Time $(Et_{C})$} is the time a service takes to respond to various types of requests.
	%is the expected delay between the time instant when a request is sent and the time when the result is obtained.
	Execution time is measured in milliseconds. Execution time can be affected by load intensity, which can be measured in terms of arrival rates (such as requests per second) or number of concurrent requests. This internal feature is a positive integer. For communities, the expected execution time of collaborating web services can be estimated as the execution time of the bottleneck service which is the service with the slowest execution time:

	\begin{equation}
		Et_{C} = \left\{
			\begin{array}{l l}
				Et_{w} & \quad \text{if $|C| = 1$}\\
				max_{w \in C}{(Et_{w})} & \quad \text{if $|C| > 1$}
			\end{array} \right.
	\end{equation}

	%\item \emph{Data Quality} The ability of a data collection to meet user requirements , defined as the proximity of a value v returned by web service to a value considered as correct. The measure of data quality is considered here as a real number in the range [0, 1], where 1 represents the most desirable score.
\end{itemize}

	We normalize the range of these features so that each feature contributes approximately proportionately to the final utility outcome value. We adopt the \emph{standardization} method. We subtract the \emph{mean} from each feature then we divide the values of each feature by its \emph{standard deviation}.

%$HI = \overline{HI}$

\subsection{External Features}\label{s:ef}

The quantitative values of quality metrics need some benchmark values to represent their goodness. In other words, without some benchmark values it would be difficult for web services to identify their goals, and performance quality at any specific value of these metrics. Therefore, we have introduced two external features for assessing web services' estimate with regard to their standing among other web services.

\begin{itemize}
  \item \emph{External Parameter 1} $(exp_1)$. is an estimate of how close the web service's \emph{execution time} is to other web services. It is the difference between a web service's \emph{execution time} metric and the maximum value of execution time of all the other web services, about which the web service has information of them. The less value means web service has better execution time compared to other web services.
	\begin{equation}\label{exp_1:f}
		exp_1 = Et_{i} - Et_{max}
	\end{equation}
	\item \emph{External Parameter 2} $(exp_2)$ is a comparison of the web service's rate of performing tasks to other competing web services. It is the difference between a web service's \emph{throughput} metric and the maximum value of throughput of all the other web services,, about which the web service has information of them. The less value means web service has better throughput compared to other web services.
	\begin{equation}\label{exp_2Lf}
		exp_2 = Th_{max} - Th_{i}
	\end{equation}
\end{itemize}

A community of web services is abstracted as a standard web service that provides services that share the same internal and external features mentioned for web services. Inside, they adopt different protocols to manage the community; however, other web services or end users can communicate with communities of web services, using standard protocols, and inner details will be encapsulated for their clients.

\subsection{Task Distribution}

Communities of web services usually employ an implementation of Contract-Net protocol for task distribution, in which services opt for incoming tasks, and receive the tasks for which they ask. However, in our model, our community members would try to distribute tasks based on their capabilities and the QoS parameters provided by the web services. We have used a slightly modified \emph{weighted fair queuing} method to distribute tasks among community members. The goal is to allocate incoming tasks to web services with a rate matching the throughput value of $Th_{ws}$. In the \emph{weighted fair queuing} method, the input flow is multiplexed along different paths; however, in our case, if the rate of incoming tasks is less than the community's total throughput $(Th_{C})$, which is the summation of throughput values of the web services in the community, some of the input tasks will be queued and served with a delay.
%Thus, the amount of tasks performed by the community is $\sum_{ws}{Th_{ws}}$ when $\sum_{ws}{Th_{ws}} \leq R_{C}$.
However, when the incoming task rate is less than the throughput of the community, the \emph{weighted fair queuing} algorithm assigns a weighted task rate of $|input Task Rate| \times \frac{Th_{ws}}{\sum_{ws}{Th_{ws}}}$ for each web service ($ws$) within the community.

While distributing tasks, the community members can verify the performance, throughput and quality of service of tasks being performed by web services. The community can assess if those web services are capable of performing the number of tasks they advertised. If for any reason, there is a decline in the quality metric or throughput, the community can consider the new values as a benchmark for future performance calculations, and penalize the suspicious web services. This way, players will have incentive to truthfully disclose their actual capabilities in order to maximize profit from the community and to avoid being penalized. In addition, the system should be dynamic enough to detect and react to web services' quality metrics variation, as over time, web service metrics may degrade or improve, changes to which the community should adjust.
% Therefore its easy for the system to encourage players to be in some sense incentive compatible in the way that they would profit best by truthfully revealing their capabilities. Also it is important to be dynamic enough to consider web services which may have their quality metrics degraded or even improved over time for any reason and be able to adjust the community with new parameters.

\subsection{Community Revenue}

Communities and web services earn revenue by performing tasks. The total gain is a function of the quality of rate of tasks being performed. The utility of a collaborative group of services $U_{C}$ is the revenue of the community. Formula \ref{u_c_normal} is an estimation of the gain of the community.

\begin{equation}\label{u_c_normal}
U_{C} = \big((\alpha \times (A_{C} - Et_{C}) - \beta \times (exp1_{C} + exp2_{C})\big) \times Th_{C}
\end{equation}

The $\alpha$ and $\beta$ values are internal and external weight coefficients. Execution time and external parameters are better when they have smaller values hence the negative coefficient for them. The result is then multiplied to the throughput value $Th_{C}$, since communities are performing tasks with $Th_{C}$ rate.

The estimation can be improved, especially in cases where the input task rate is high and services are experiencing high task loads. The \emph{weighted fair queuing} method of task distribution would distribute tasks based on the individual throughput $(Th_{w})$ value of services within community. The services with higher throughput will affect the overall utility of the community more because they would take on proportionately more tasks.

\begin{equation}\label{u_c_load}
\begin{split}
U_{C} = \sum_{ws \in C}&\bigg(\big(\alpha \times (A_{ws} - Et_{ws}) \\
        & - \beta \times (exp1_{ws} + exp2_{ws})\big) \times Th_{ws}\bigg)
\end{split}
\end{equation}

\section{The Decision Making Mechanism}\label{s:model}

In this section, we describe our data extraction process and the methodology used to equip web services with a decision making mechanism. In this methodology, we first present the data extraction and engineering process and then we evaluate the decision making mechanism for web services in community settings. Figure \ref{fig_steps} summarizes the steps performed in DDM from the input data to generation of decision making profiles for web services.

\begin{figure}%[!t]
\centerline{\includegraphics[width=1.75in]{figures/steps.eps}}
\caption{A summary of DDM decision profile generation steps}
\label{fig_steps}
\end{figure}

\subsection{Data Extraction and Solution Engineering}\label{ss:learningdata}

\subsubsection{Input Web Services Data}\label{sss:webservices}

%We used a data set extracted from running web services. In this data set,
Each web service is associated with a number of features that reflect its functionality. These web services operate in an online environment and are continuously assigned tasks to handle. Sometimes, they pass tasks around, distributing sub-tasks among themselves and handling tasks in a collaborative effort. To engage in communities, we add additional features that reflect web services' utility as a result of joining other web services to form a community. The additional features are computed based on two assumptions that we adopt for a community to be formed. Our raw data set provides us with 64 different time slots of extracted features for each web service. Using this data we can build a synthetic data set that contains features of a large number of web services in a sequence of time. Therefore, we train a learning model that adopts the trend of joining a community and use the model to predict/find the appropriate community for other web services. To train our model, we build a decision tree as a benchmark for our decision making mechanism. The decision tree is built by training the real data obtained from operating web services and extracting features related to their performance, either alone or as part of a joint effort with other web services. As a result, we are able to classify web services as either single or joined and propose the communities that are best fitted to the web services' requirements.

\subsubsection{Filtered Data}\label{sss:filtereddata}

We have adopted the web service data set provided by \cite{10.1109/ISSRE.2011.17}. The raw data provides real-world QoS evaluation results from several users on 5,825 web services over 64 different time frames\footnote{http://www.wsdream.net/}. In this data set, each web service is associated with a number of features that reflect its functionality. By summarizing the data provided for each web service over different time slots, we have acquired three quality features for each individual web service. Their quality features are the three internal features introduced in Section \ref{s:if}: \emph{throughput}, \emph{availability} and \emph{execution time}.
%To engage in communities, we add two additional external features that reflect web services estimation of its quality metrics compared to other web services.
Therefore, we have generated an array of web services with three distinct features.

\begin{figure}%[!t]
\centerline{\includegraphics[width=3.25in]{figures/cfvs.eps}}
\caption{Communities with different properties of web services actively looking for other communities to collaborate with, they send collaboration requests to communities they find the collabration  increasing the gain}
\label{fig_community}
\end{figure}

Now, we want to generate a set of feature vectors representing the communities of web services for training purposes. For communities, we add two additional external features, as discussed in Section \ref{s:ef} that reflect the community's estimation of its quality metrics compared to other web services. Therefore, we formulate a \emph{Community Feature Vector (CFV)} as $CFV_{<C>} = \{f_1,...f_5\}$ having a community of $k$ web services. $C = {ws_1,...ws_k}$. The features $f_1$ through $f_5$ represent the \emph{execution time}, \emph{throughput}, \emph{availability} and the \emph{external parameters 1 and 2} correspondingly. A set of communities, with their feature vectors and utilities evaluated, provides our algorithm with a raw training data set. We call this set of communities the \emph{template vector}, and the set of feature vectors associated with the \emph{template vector} is referred to as the \emph{community feature vector set (CFVS)}. Figure \ref{fig_community} depicts communities looking to form collaborate communities and join other communities in order to improve their utility gain.

\subsubsection{Feature Engineering}\label{sss:feng}
Let $CFVS = \{C_1, C_2,..., C_N\}$ be the community feature vector set with $N$ communities. Based on the $CFVS$ set, we create an $|N \times N|$ gain matrix for $t$ different time slots. Each entry of $gain_{n,m}^{t}$corresponds to a utility gain of web services in $C_n$, when $C_n$ joins $C_m$, based on community and web service data available for time $t$. All of these values have to be evaluated by implementing the community with members of $ws_i \in C_n$ and $ws_j \in C_m$ and evaluating its utility (see Equation \ref{u_c_load}). Therefore $gain_{n,m}^{t} = U_{C_n \cup C_m}^{t} - U_{C_{n}}^{t}$. Evaluating the utility gain for all entries of the $gain$ matrix can be a costly process with a very big number of $N$, size of the feature vector $|N|$ should be chosen carefully.

\begin{table*}[ht]
\caption{An example of $gain$ matrix for 3 different communities and their combinations} % title of Table
\centering % used for centering table
{\renewcommand{\arraystretch}{1.2}
\begin{tabular}{c|c c c c c c} % centered columns (4 columns)
\hline\hline %inserts double horizontal lines
 & \textless348\textgreater & \textless1934\textgreater & \textless2117\textgreater & \textless348, 1934\textgreater & \textless1934, 2117\textgreater & \textless348, 1934, 2117\textgreater \\ [0.5ex] % inserts table
%heading
\hline % inserts single horizontal line
\textless348\textgreater & - & 0.282708 & 1.027081 & 0.282708 & 18.027081 & 18.027081 \\
\textless1934\textgreater & -2.637483 & - & 6.969072 & -2.637483 & 5.509583 & 4.387725 \\
\textless2117\textgreater & 5.027081 & 2.969072 & - & 5.509583 & 2.969072 & 5.509583 \\
\textless348, 1934\textgreater & 0.0 & 0.0 & -3.851432 & - & -3.851432 & -3.851432 \\
\textless1934, 2117\textgreater & 2.969072 & 0.0 & 0.0 & 2.969072 & - & 2.969072 \\
\textless348, 1934, 2117\textgreater & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & - \\ [1ex] % [1ex] adds vertical space
\hline %inserts single line
\end{tabular}
}
\label{table:nonlin} % is used to refer this table in the text
\end{table*}

Now, we let our set of communities in the $CFVS$ set, within $|T|$ time frame iterations, choose the the best communities to join. Each community is provided with the corresponding row of data from the $gain$ matrix. Basically, $C_i$ is provided with the data in row $i$ of matrix $gain_{N,N}$, which is all of the possible utility values it can gain by joining different communities. By ordering the list, each community is utilized with an ordered set of preferences over other communities it likes to join. We define $>_{i}^t$ as the preference order of community $i$ at time $t$.

Let $P_{C_i}^t = \{C_1 >_{i}^t C_2 >_{i}^t .... >_{i}^t C_n\}$ be an ordered set of preferences for community $C_i$ at time $t$. Based on this ordered set, we define $K^t(C_i, k)$ as a set of the $k$ most preferred communities of community $i$ at time $t$.
%\subsubsection{feature vector generation}\label{sss:fvg}
\begin{equation}\label{h_t_pref_top}
\begin{split}				
K^t(C_i, k) = &\Big\{C_x | C_x >_{i}^t C_y; \forall C_y \in CFVS\ \\
				      &\wedge C_y \notin \{K^t(i, 1),K^t(i, 2),...,K^t(i, k-1)\} \Big\}				
\end{split}
\end{equation}
Based on $K^t(C_i, k)$, we define a set of communities for $C_i$ in which they are the k most preferred communities for community $i$ and also community $i$ belongs to the k most preferred communities out of all of the communities in set $K^t(C_i, k)$. This basically implies the preference is both-sided.
\begin{equation}\label{l_t_top_both}
\begin{split}	
L^t(C_i,k) = &\Big\{C_j | C_j \in K^t(C_i, k) \\
             &\wedge C_i \in \bigcup_{C_x \in K^t(C_j, k)}K^t(C_x, k)\Big\}
\end{split}
\end{equation}

Table \ref{table:nonlin} illustrates an example of a $gain$ matrix for 3 different communities and their combinations. Each row shows the gain the community can achieve by collaborating with other 5 communities. In this example for community $\textless 348 \textgreater$ we have: \\
$K(\textless 348 \textgreater, 1) = \{\textless1934, 2117\textgreater\}$ and \\
$K(\textless 348 \textgreater, 2) = \{\textless1934, 2117\textgreater, \textless1934\textgreater\}$ \\
Since $\textless 348 \textgreater$ is best preferred community of $\textless 1934, 2117 \textgreater$ therefore $L(\textless 348 \textgreater, 1)$ is not empty and contains the community $\textless1934, 2117\textgreater$.

Using the $gain$ matrix and the mentioned preference ordering relations, we are able to build a decision tree where the list of possible communities to join and their expected utilities are set.
% as well as the joined events that took place in different time slots.
In addition to best choices, web services have access to other ordered choices and can look for the second best or third best if their first try is rejected by the target community. The consequences of each try level is analyzed in more detail in the following section, in which we launch experiments and investigate the effectiveness of the use of a decision tree with different decision layers in joining other communities and enhancing the overall utility.

\subsection{Decision Profile Generation}\label{ss:learningmodel}

Our goal is to create a decision making profile for each community in the $CFVS$ set. We are creating an environment where they can experience the outcomes of different strategies. The result will be a decision tree of the feasible and utility increasing moves over time.

We let communities pick the best communities for maximizing their utilities over different time frames. At time $t = 1$, we let each community in the $CFVS$ set choose the best community, which is a single community in set $C_j = K^t(C_i, 1)$. If community $C_j$ also has the highest preference to join $C_i$ meaning the $L^t(C_i, k=1)$ set is not empty, they would join together. Having set $k = 1$ is a very strict and hardly satisfiable condition. In order to relax the requirement, the value of $k$ is being increased by a rate $r$ proportional to time slot $t$. Basically, $k = 1 + |r \times t|$. On early steps of simulation web services and communities are more strict but as time goes on, we let them choose second and then third best options too. However, increasing $k$ quickly increases the time complexity of $K^t(C_i, k)$ by $O(k^2 \times n.log(n))$.

When communities $C_i$ and $C_j$ are in each other's top $k$ preference set, their combination is added to the list of possible communities that can join others at time $t+1$. Also, for each community $C_i$ in our initial $CFVS$ set, we maintain a tree with the community $C_i$ as its root. Its children are all of the communities that $C_i$ decided to join. As the scenario progresses over time, the merged communities may decide to join other communities. When community $C_i$ and $C_j$ decide to join each other and create community $C_k$, the new community $C_k$ will be added as a child to both $C_i$ and $C_j$ nodes. At the end of the process, each community is utilized with a tree of set of possible combination of communities it can join. Algorithm \ref{algo:dectree} illustrates the DDM tree creation procedure as pseudo-code.

Having created $|n|$ trees, for each community $C_i$, our communities are utilized with the different possible paths they can take to maximize their utilities. Using a distance function, communities and web services can utilize these results and find the $C_i$ that closely resembles their parameters. They will be utilized with a tree that has the strategies the community can use to join our communities which will be best possible path to take and communities to join while most likely being accepted by other communities since its a two sided preference. The algorithm below describes the tree creation process.

\begin{algorithm}
\DontPrintSemicolon
\KwIn{$\langle r, gain^t_{n,n}, CFVS \rangle$ learning rate $r$, $|N \times N \times T|$ gain matrix, community feature vector}
\KwOut{A set of \emph{root} nodes of the decision trees}
$k \gets 1$\;
$nodes[N] \gets$ initialize $N$ tree nodes representing each community in CFVS\;
\For{$t \gets 1$ \textbf{to} $T$} {
	$k \gets 1 + round (r \times t)$\;
  \For{all $C_i \in CFVS$} {
	  \For{all $C_j \in L^t(C_i, k)$} {
      % The "l" before the If makes it so it does not expand to a second line
      \If{$C_i \in L^t(C_j, k)$}{
        $C_k \gets C_i \cup C_j$\;
				add $C_k$ to $CFVS$ set\;
				initiate $node_k$, representing $C_k$\;
				$nodes_i.addChild (node_k)$\;
				$nodes_j.addChild (node_k)$\;
      }
%      \Else{
%        $j \gets j + 1$\;
%      }			
		}
  }
%  $i \gets i + 1$\;
}
\Return{nodes}\;
\caption{{\sc DDM Decision Tree Algorithm}}
\label{algo:dectree}
\end{algorithm}

\textbf{Complexity.} Here we analyze the computational complexity of the DDM decision tree creation algorithm on each time iteration $t$. Computing top $k$ preferred communities for $C_i$ in $L^t(C_i, k)$ requires an $O(n.log(n))$ sort time. $C_i$ has a result set of size $|k|$, and for each of those $k$ communities we need to check against their $k$ top preferred communities. Considering we already have the list sorted, line 6 in algorithm runs $O(k \times n.log(n))$ times. Doing this search for all communities in $|CFVS|$ set $n$ times (line 5), the overall order of complexity of the algorithm in regards with $n$ and $k$ is: $O(k^2 \times n^2.log(n))$.

\section{Experiments}\label{s:experiments}
We implemented DDM in Java\footnote{Source code of implementation and data is available at: \emph{https://github.com/Marooned202/DDM}}. We have extracted the set of features for 4532 web services in 64 different time slots through a data set provided by \cite{10.1109/ISSRE.2011.17}. By choosing 86 web services out of these sets and choosing a subset of all possible combinations of sizes 2, 3, and 4 of these 86 web services, we have chosen 10,000 communities and evaluated the feature vectors and utilities they can have in 64 different time slots. This provides us with the initial training feature set of size $|CFVS| = 10,000$ communities. Based on Equation \ref{u_c_load} the utility of these communities are estimated, and the gain matrix $|10000 \times~ 10000|$ of all possible ways of merging these 10,000 communities, is generated\footnote {The template vector and gain matrix generated are available at \emph{https://github.com/Marooned202/DDM/tree/master/wsds/data/run}}, and each community has the ordered preference among other communities known in the set.

We let communities and web services adopt their strategies based on our $DDM$ decision making mechanism. Based on the decisions adopted, each community will generate a decision tree profile. We let DDM run four times with different $r$ rates of 0.05, 0.07, 0.10 and 0.20. With the slow rate of $r = 0.05$, we increase $k$ in Equation \ref{u_c_normal} for every 20 time frames, which will happen only three times in our 64-step experiment. In the case of $r = 0.20$, k increases much faster, at a rate of once every 5 time frames, which increases the complexity of the $L^t(C_i,k)$ search for each community in the $CFVS$ set.

Table \ref{table:valueandgain} depicts the average utility gain value of the communities in each of the four runs. The utility gain is the increase of utility the communities gain by cooperating and joining other communities. The utility gain ratio is the ratio of their final utility over initial utility. Comparing the different search rates, we can see increasing $r$ to value of $0.10$ results in significant performance boost, however the higher rates of $r$ are not increasing the chance of finding better collaborative groups for our communities.

\begin{table}[ht]
\caption{Utility gain of web services after making collaborative groups based on DDM algorithm with different $r$ rates} % title of Table
\centering % used for centering table
{\renewcommand{\arraystretch}{1.2}
\begin{tabular}{c|c|c} % centered columns (4 columns)
\hline\hline %inserts double horizontal lines
Search Rate $r$ & Utility Gain Value & Utility Gain Ratio \\ [0.5ex] % inserts table
%heading
\hline % inserts single horizontal line
r=0.20 & 176.1499 & 6.9690 \\
r=0.10 & 174.6541 & 6.9182 \\
r=0.07 & 159.9462 & 6.2834 \\
r=0.05 & 136.0768 & 5.1032 \\ [1ex] % [1ex] adds vertical space
\hline %inserts single line
\end{tabular}
}
\label{table:valueandgain} % is used to refer this table in the text
\end{table}

%\begin{figure}%[!t]
%\centering
%\includegraphics[width=3.5in]{figures/utility_gain_r.eps}
%\caption{DDM Utility Gain Value}
%\label{utility_gain_value}
%\end{figure}
%
%\begin{figure}%[!t]
%\centering
%\includegraphics[width=3.5in]{figures/utility_ratio_r.eps}
%\caption{DDM Utility Gain Ratio}
%\label{utility_gain_ratio}
%\end{figure}

We utilize the same communities and web services with a simple rational decision making mechanism, in which communities will choose to join a community if it increases utility by any amount. We call it the $rational$ method, in the sense that it is just a rational method and is not aiming to be optimal. We chose 10 random web services and compared the results with web services which adopted our DDM model. Figure \ref{utility_gain_mlisa_and_rational} shows the comparison of the end result of utility gain values. In 18 out of 40 tries, \emph{rational} agents were not able to improve their utility at all because the communities they chose rejected their request, most likely because they would not have increased the utility of the other communities if they had joined them. The results show that a long-term strategic decision mechanism is needed to satisfy all of the services within communities. Figure \ref{utility_gain_mlisa_and_rational_ratio} shows the same results in terms of ratio of utility gain.
